Simulating cosmic variance in Weak Lensing: effects on parameter inferences

### 2015-11-20 Paper/paper.tex
second draft

\begin{abstract}
Constraining cosmology using Weak Gravitational Lensing consists in comparing measured features with simulated ones. An accurate estimate of the feature covariance matrix is essential to obtain accurate parameter confidence intervals. When this covariance matrix $\bb{C}$ is measured from simulations, an important question to ask is how big the simulation set should be for an accurate estimation of $\bb{C}$. We construct different mock ensembles with $N_r$ realizations of the same shear field, based on $N_s<N_r$ independent $N$--body simulations. Using the shear--shear power spectrum as a summary statistic, we find that a single $N$--body simulation can be recycled $O(10^6)$ times to produce an ensemble of $N_r=10^5$ shear realizations that are mutually independent. We also find that, for a range of summary statistics that include shear--shear power spectra measured at different scales and converge peak counts, parameter constraints degrade faster than the $O(N_b)$ Gaussian prediction by \citep{DodelsonSchneider13}, where $N_b$ is the length of the feature vector.       
\end{abstract}

@@

\subsection{Shear field simulations}
\label{shearsim}
%
In this paragraph we describe how we constructed our shear field ensembles. Background galaxies at redshift $z_s$ are lensed by large scale structure between $z=0$ and $z_s$. The shape distortions due to the cosmic shear $\pmb{\gamma}$ can be computed in terms of the dark matter gravitational potential $\Phi(\bb{x},z)$. Because the evolution of $\Phi$ in redshift is non--linear, its evolution needs to be computed with numerical simulations. We make use of the public code Gadget2 \citep{Gadget2}, with which we run a sequence of dark--matter--only $N$--body simulations that track the evolution of the density fluctuations in the large scale structure of the universe. We assume a standard $\Lambda$CDM framework with $(\Omega_m,\Omega_\Lambda,h,w,\sigma_8,n_s)=(0.26,0.74,0.72,-1,0.8,0.96)$ and we fix the comoving size of the simulation box to $240\mathrm{Mpc}/h$. We fill the box with $512^3$ particles, which correspond to a mass resolution of $10^{10}M_\odot$. We assume a uniform galaxy distribution at a constant redshift $z_s=2$ (at which the simulation box has an angular size of $\theta_{box}=3.5\mathrm{deg}$) and we disctretize the mass distribution between $z_s$ and the observer at $z=0$ with a sequence of 60 two dimensional lenses of thickness $80\mathrm{Mpc}/h$. The surface density on each lens is computed by projecting the three dimensional density measured from Gadget2 snapshots. We then apply the multi--lens--plane algorithm (see \citep{RayTracingHartlap,RayTracingJain} for example) to trace the deflections of $2048^2$ light rays arranged on a square grid of total size $\theta_{box}$, from $z=0$ to $z_s$. The light deflection calculations have been performed with our implementation of multi--lens--plane algorithm, which is part of the LensTools computing package \citep{LensTools} (released under the MIT license). Different realizations $r$ of the same shear field $\pmb{\gamma}_r(\pmb{\theta})$ can be generated with different choices of the lenses that lay between the observer and $z_s$. The randomization procedure we adopt is the following:

@@

\begin{equation}
\label{covest}
\bbh{C} = \frac{1}{N_r-1}\sum_{r=1}^{N_r} (\bbh{d}_r - \bar{\bb{d}}) (\bbh{d}_r - \bar{\bb{d}})^T
\end{equation}
%
Assuming a normal feature likelihood, together with a flat prior on the parameter space, the parameter posterior distribution $\mathcal{L}(\bb{p}\vert\bbh{d})$ can be obtained using Bayes theorem
\begin{equation}
\label{posteriorbayes}
\mathcal{L}(\bb{p}\vert\bbh{d}) = \mathcal{N}_\mathcal{L}\exp{\left[-\frac{1}{2}(\bbh{d}-\bb{d}(\bb{p}))^T\bb{C}^{-1}(\bbh{d}-\bb{d}(\bb{p}))\right]}
\end{equation}
%
For the sake of simplicity, we approximate the posterior as a Gaussian around its maximum. This is easily done approximating the simulated feature at first order around a point $\bb{p}_0$ (that ideally is the maximum of (\ref{posteriorbayes}))
\begin{equation}
\bb{d}(\bb{p}) \approx \bb{d}_0 + \bb{d}_0^\prime(\bb{p}-\bb{p}_0) 
\end{equation}
%
From which we can build the estimators for the parameters expectation value $\bbh{p}$ and covariance $\h{\Sigma}_\bb{p}$, given the observation $\bbh{d}$, as follows:
%
\begin{equation}
\label{estimatormean}
\bbh{p} = \bb{p}_0 + \bbh{T}(\bbh{d}-\bb{d}_0)
\end{equation}

\begin{equation}
\label{estimatorcovariance}
\h{\Sigma}_\bb{p} = \bbh{T}(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\bbh{T}^T
\end{equation}

\begin{equation}
\bbh{T} = (\bb{d}_0'^T\bbh{C}^{-1}\bb{d}_0')^{-1}\bb{d}_0'^T\bbh{C}^{-1}
\end{equation}
%
Where, for simplicity we assumed $\langle\bbh{d}\rangle=\bb{d}_0$, so that $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$. The noise in the covariance estimator (\ref{covest}) propagates all the way to the posterior (\ref{posteriorbayes}), the parameter estimate (\ref{estimatormean}) and its variance (\ref{estimatorcovariance}), through the transformation matrix $\bbh{T}$. Following \citep{Taylor12} we can say that, if the feature vector $\bbh{d}$ from which $\bbh{C}$ is estimated is drawn from a multivariate Gaussian distribution, the expectation values of the fluctuations of its inverse $\bbh{\Psi}$ can be computed analytically as:

\begin{equation}
\langle\delta\h{\Psi}_{ij}\delta\h{\Psi}_{kl}\rangle = A\Psi_{ij}\Psi_{kl} + B(\Psi_{ik}\Psi_{jl} + \Psi_{il}\Psi_{jk})
\end{equation}
%
where $A,B$ are constants that depend on $N_r,N_b$ and have the asymptotic behavior $A\propto1/N_r^2,B\propto1/N_r$. \citep{DodelsonSchneider13} calculated the effect of the $\bbh{\Psi}$ fluctuations on the parameter constraints and found that the parameter variance estimator (\ref{estimatorcovariance}) is biased

\begin{equation}
\label{dodelsonscaling}
\langle\bbh{\Sigma}_\bb{p}\rangle = \left[1+B(N_b-N_p)\right]\bb{\Sigma}_\bb{p}
\end{equation} 
%
With the $\bbh{d}$ Gaussianity assumption, the coefficient $B$ can be calculated exactly as \citep{DodelsonSchneider13,Taylor12}:
\begin{equation}
B = \frac{N_r-N_b-2}{(N_r-N_b-1)(N_r-N_b-4)}
\end{equation}
%
In the limit $N_r\gg N_b$ this leads to a degradation of order $(N_b-N_p)/N_r$ for the variance of each estimated parameter $\sigma^2_p=\Sigma_{pp}$. We test the scaling relation
\begin{equation}
\label{ourscaling}
\langle\h{\sigma}_p^2\rangle = \sigma^2_p(N_s)\left(1+\frac{D}{N_r}\right)
\end{equation}
%
on our simulations. We call $D$ the \textit{effective dimensionality} of the feature space which, under Gaussianity assumptions, is equal to $N_b-N_p$. We compute the expectation values of $\h{\sigma}^2_p$ (equations (\ref{estimatorcovariance}),(\ref{ourscaling})) averaging over 100 random resamplings of our shear ensembles. For the true feature covariance matrix $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$ we use the estimated covariance from a grand ensemble built with the union of all the ensembles with different $N_s$. 
The true parameter variance $\sigma^2_p(N_s)$ in principle depends on the number of independent $N$--body simulations $N_s$, which appears in the randomization procedure we describe in \S~\ref{shearsim}. The reason for this can be found in that, if $N_s$ is not big enough, the different shear realizations cannot be all independent, and hence the true variance $\sigma^2_p(N_s\rightarrow\infty)$ cannot be recovered in the limit $N_r\rightarrow\infty$. In particular there could be some residual error in the expectation value of (\ref{estimatorcovariance}) even if $N_r$ is big. In the next paragraph we outline our main findings.

@@

\begin{figure*}
\includegraphics[scale=0.4]{Figures/ps_pdf.eps}
\caption{PDF of the $\kappa$ power spectrum,$\mathcal{L}(P_l^{\kappa\kappa})$, at four selected multipoles $l=115,344,1349,5283$, for different shear ensembles constructed with $N_s=$1 (black), 2 (blue), 5 (green), 50 (red), 100 (purple).}
\label{ps_pdf}
\end{figure*}

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_nr.eps}
\caption{Expectation value of the variance on $w$, $\langle\h{\sigma}^2_w\rangle$, as a function of the number of mock realizations $N_r$ used in the feature covariance estimator (\ref{covest}). Results for different mock ensembles generated with $N_s=$1 (black), 2 (blue), 5 (green), 50 (red), 100 (purple) are shown. The figure shows both the measured trends (solid lines) and their best fits according to equation (\ref{dodelsonscaling}). }
\label{wvar_nr}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_ns.eps}
\caption{Expectation value of the variance on $w$, $\langle\h{\sigma}^2_w\rangle$, varying the mock ensemble used to compute the covariance estimator (\ref{covest}). We factored out the $(1+D/N_r)$ dependence, showing only the dependence of $\sigma_0^2$ on $D$. Different choices of $N_r=$100 (black), 300 (blue), 500 (green), 700 (red), 900 (purple) are shown for comparison.}
\label{wvar_ns}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\toprule
\textbf{Feature} &  \textbf{Specifications} & $N_b$ &  \textbf{Symbol} & \textbf{Color} \\ \hline \hline
\midrule
Power Spectrum, log binning  & $l \in [100,800] $ & 8 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [1000,6000] $ & 7 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [100,6000] $ & 15 & $\times$ & black  \\
Power Spectrum, linear binning  & $l \in [100,2000] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [2500,4500] $ & 15 & \textcolor{red}{$\times$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $l \in [100,4500] $ & 30 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [100,6000] $ & 39 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\ \hline
Low peaks  & $\kappa_0 \in [-0.06,0.09] $ & 15 & \textcolor{OliveGreen}{$+$} & \textcolor{OliveGreen}{green}  \\ 
Intermediate peaks  & $\kappa_0 \in [0.1,0.27] $ & 15 & \textcolor{OliveGreen}{$\bigstar$} & \textcolor{OliveGreen}{green}  \\ 
High peaks  & $\kappa_0 \in [0.28,0.45] $ & 15 & \textcolor{OliveGreen}{$\diamond$} & \textcolor{OliveGreen}{green}  \\
Low+Intermediate peaks  & $\kappa_0 \in [-0.06,0.27] $ & 30 & \textcolor{OliveGreen}{$\times$} & \textcolor{OliveGreen}{green}  \\
Intermediate+High peaks  & $\kappa_0 \in [0.1,0.45] $ & 30 & \textcolor{OliveGreen}{$\bullet$} & \textcolor{OliveGreen}{green}  \\
All peaks  & $\kappa_0 \in [-0.06,0.45] $ & 45 & \textcolor{OliveGreen}{$\blacksquare$} & \textcolor{OliveGreen}{green}  \\ \hline
\bottomrule
\end{tabular}
\end{center}
\caption{Effective dimensionality of the power spectrum feature space $D$ computed fitting the measured $w$ variance with the scaling relation (\ref{dodelsonscaling}). We show both cases in which we consider the full covariance matrix and the case in which we artificially zero out all the off diagonal elements.}
\label{dimtable}
\end{table*}

\begin{figure}
\includegraphics[scale=0.35]{Figures/effective_nb.eps}
\caption{}
\label{effectivenb}
\end{figure}

If the feature we consider is the shear-shear power spectrum, its derivatives $\bb{d}_0'$ with respect to the cosmological parameters can be evaluated by existing analytical codes. For this purpose we use the analytical code NICAEA (see \citep{Nicaea}). In this section we outline the main results of this work. We show the qualitative behavior of the $P^{\kappa\kappa}$ probability distribution function (PDF) in ensembles built with different $N_s$. We also show how the forecasts on the Dark Energy equation of state, $w$, depend on $N_s$ and $N_r$. Figure \ref{ps_pdf} shows the power spectrum PDF at four selected multipoles, while Figure \ref{ps_var} shows the variance in each multipole. This result is compared to the one that we would obtain assuming that the convergence is a Gaussian random field
\begin{equation}
\label{gaussianvar}
\mathrm{Var}(P^{\kappa\kappa}(l)) = \frac{P^{\kappa\kappa,2}(l)}{N(l)}
\end{equation}
%
where $N(l)$ is the number of modes used to estimate the power spectrum at $l$, which depends on the mock size $\theta_{box}$ and on the width of the $N_b$ multipole bands. Figure \ref{wvar_nr} and Table \ref{dimtable} show the scaling of the expectation value of $\h{\sigma}_w^2$ with the number of mocks $N_r$ and compare it with the known result obtained in \citep{DodelsonSchneider13}. Figure \ref{wvar_ns} shows the effect of varying $N_s$ on the $w$ constraint.

@@

In this section we discuss our main findings. Figure \ref{ps_pdf} shows that, although different choices of $N_s$ do not seem to affect the power spectrum PDF at large scales (top two panels), there are some qualitative differences on small scales (bottom two panels), on which a shear ensemble built with a small $N_s$ might not have the same statistical behavior of ensembles built with large $N_s$. Figure \ref{ps_var} shows the variance of convergence power spectrum computed from different ensembles, in units of the Gaussian expectation. We find that, even with $N_s=1$, our results are in good agreement with the ones obtained by \citep{Sato12}, which used $N_s=400$. This fact by itself is not sufficient to conclude that $N_s$ does not have an effect on parameter inferences, since these depend on the cross band covariances.

In Figure \ref{wvar_nr} we examine how the $w$ constraint depends on the number of mocks used to estimate the covariance, and we find good agreement with \citep{DodelsonSchneider13}. We determine the effective dimensionality of the feature space, $D$, with a linear regression using equation (\ref{ourscaling}) and we find that this depends strongly on how the noise is distributed across the covariance matrix estimator $\bbh{C}$. In particular we find that when every element in the covariance is subject to noise $D$ is close to the number of bands used $N_b=15$, but when we artificially zero out the off diagonal elements of $\bbh{C}$, the effective dimensionality drops to $D\approx2.5$. Finally, Figure \ref{wvar_ns} shows the dependence of the dark energy constraint with $N_s$ and $D$. We find that, in the range $N_s\in[1,200]$ the $w$ inferred variance has sub--percent fluctuations for both the diagonal and full $\bbh{C}$ cases, while there is a $\sim$10\% change in $\sigma^2_0(N_s,w)$ when the effective dimensionality $D$ is different. Finally, the different lines in Figure \ref{wvar_ns} show that equation (\ref{ourscaling}) is a good model for the variance scaling.

>>>>

\begin{abstract}
Constraining cosmology using Weak Gravitational Lensing consists in comparing measured features with simulated ones. An accurate estimate of the feature covariance matrix is essential to obtain accurate parameter confidence intervals. When this covariance matrix $\bb{C}$ is measured from simulations, an important question to ask is how big the simulation set should be for an accurate estimation of $\bb{C}$. We construct different mock ensembles with $N_r$ realizations of the same shear field, based on $N_s<N_r$ independent $N$--body simulations. Using the shear--shear power spectrum as a summary statistic, we find that a single $N$--body simulation can be recycled a large number of times to produce an ensemble of $N_r=10^5$ shear realizations that are mutually independent. We also find that, for a range of summary statistics that include shear--shear power spectra measured at different scales and convergence peak counts, parameter constraints degrade faster than the $O(N_b)$ Gaussian prediction by \citep{DodelsonSchneider13}, where $N_b$ is the length of the feature vector.       
\end{abstract}

@@

\subsection{Shear field simulations}
\label{shearsim}
%
In this paragraph we describe how we constructed our shear field ensembles. Background galaxies at redshift $z_s$ are lensed by large scale structure between $z=0$ and $z_s$. The shape distortions due to the cosmic shear $\pmb{\gamma}$ can be computed in terms of the dark matter gravitational potential $\Phi(\bb{x},z)$. Because the evolution of $\Phi$ in redshift is non--linear, its evolution needs to be computed with numerical simulations. We make use of the public code \texttt{Gadget2} \citep{Gadget2}, with which we run a sequence of dark--matter--only $N$--body simulations that track the evolution of the density fluctuations in the large scale structure of the universe. We assume a standard $\Lambda$CDM framework with $(\Omega_m,\Omega_\Lambda,h,w,\sigma_8,n_s)=(0.26,0.74,0.72,-1,0.8,0.96)$ and we fix the comoving size of the simulation box to $240\mathrm{Mpc}/h$. We fill the box with $512^3$ particles, which correspond to a mass resolution of $10^{10}M_\odot$. We assume a uniform galaxy distribution at a constant redshift $z_s=2$ (at which the simulation box has an angular size of $\theta_{box}=3.5\mathrm{deg}$) and we disctretize the mass distribution between $z_s$ and the observer at $z=0$ with a sequence of 46 two dimensional lenses of thickness $80\mathrm{Mpc}/h$. The surface density on each lens is computed by projecting the three dimensional density measured from \texttt{Gadget2} snapshots. We then apply the multi--lens--plane algorithm (see \citep{RayTracingHartlap,RayTracingJain} for example) to trace the deflections of $2048^2$ light rays arranged on a square grid of total size $\theta_{box}$, from $z=0$ to $z_s$. The light deflection calculations have been performed with our implementation of multi--lens--plane algorithm, which is part of the \texttt{LensTools} computing package \citep{LensTools} (released under the MIT license). Different realizations $r$ of the same shear field $\pmb{\gamma}_r(\pmb{\theta})$ can be generated with different choices of the lenses that lay between the observer and $z_s$. The randomization procedure we adopt is the following:

@@

\begin{equation}
\label{covest}
\bbh{C} = \frac{1}{N_r-1}\sum_{r=1}^{N_r} (\bbh{d}_r - \bar{\bb{d}}) (\bbh{d}_r - \bar{\bb{d}})^T
\end{equation}
%
Assuming a normal feature likelihood, together with a flat prior on the parameter space, the parameter posterior distribution $\mathcal{L}(\bb{p}\vert\bbh{d})$ can be obtained using Bayes theorem
\begin{equation}
\label{posteriorbayes}
\mathcal{L}(\bb{p}\vert\bbh{d}) = \mathcal{N}_\mathcal{L}\exp{\left[-\frac{1}{2}(\bbh{d}-\bb{d}(\bb{p}))^T\bb{C}^{-1}(\bbh{d}-\bb{d}(\bb{p}))\right]}
\end{equation}
%
For the sake of simplicity, we approximate the posterior as a Gaussian around its maximum. This is easily done approximating the simulated feature at first order around a point $\bb{p}_0$ (that ideally is the maximum of (\ref{posteriorbayes}))
\begin{equation}
\bb{d}(\bb{p}) \approx \bb{d}_0 + \bb{d}_0^\prime(\bb{p}-\bb{p}_0) 
\end{equation}
%
To measure the feature derivatives $\bb{d}'_0$ with respect to cosmology, we make use of the public code code NICAEA (see \citep{Nicaea}) for the power spectrum, and we use an independent simulation set (that contains mocks at different combinations of $(\Omega_m,w,\sigma_8)$) for the peak counts.
We can build the estimator for the posterior location parameter $\bbh{p}$, given the observation $\bbh{d}$, as follows:
%
\begin{equation}
\label{estimatormean}
\bbh{p} = \bb{p}_0 + \bbh{T}(\bbh{d}-\bb{d}_0)
\end{equation}

\begin{equation}
\bbh{T} = (\bb{d}_0'^T\bbh{C}^{-1}\bb{d}_0')^{-1}\bb{d}_0'^T\bbh{C}^{-1}
\end{equation}
%
Defining the \textit{precision matrix} $\bbh{\Psi}=\bbh{C}^{-1}$, we can express the estimator of the parameter covariance as

\begin{equation}
\label{estimatorcovariance}
\h{\Sigma}_\bb{p} = \bbh{F}^{-1}\bb{d}_0'^T\bbh{\Psi}\bb{C}\bbh{\Psi}\bb{d}_0'\bbh{F}^{-1}
\end{equation}

\begin{equation}
\label{estimatorfisher}
\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'
\end{equation}
%
Where we introduced the Fisher matrix estimator $\bbh{F} = \bb{d}_0'^T\bbh{\Psi}\bb{d}_0'$ and, for simplicity, we assumed $\langle\bbh{d}\rangle=\bb{d}_0$, so that $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$. 
The noise in the covariance estimator (\ref{covest}) and its inverse $\bbh{\Psi}$ propagates all the way to the posterior (\ref{posteriorbayes}), the parameter estimate (\ref{estimatormean}) and its variance (\ref{estimatorcovariance}). Following \citep{Taylor12} we can say that, if the feature vector $\bbh{d}$ from which $\bbh{C}$ is estimated is drawn from a multivariate Gaussian distribution, the precision matrix estimator $\bbh{\Psi}$ is biased
\begin{equation}
\label{psibias}
\langle\bbh{\Psi}\rangle = \frac{N_r-1}{N_r-N_b-2}\bb{\Psi}
\end{equation}
%
and, once the bias is corrected, the residual fluctuations in $\bbh{\Psi}$ are correlated as follows
\begin{equation}
\label{psifluctuations}
\langle\delta\h{\Psi}_{ij}\delta\h{\Psi}_{kl}\rangle = A\Psi_{ij}\Psi_{kl} + B(\Psi_{ik}\Psi_{jl} + \Psi_{il}\Psi_{jk})
\end{equation}
%
where $A,B$ are constants that depend on $N_r,N_b$ and have the asymptotic behavior $A\propto1/N_r^2,B\propto1/N_r$. We want to estimate the effect of these fluctuations on the parameter constraints $\Sigma_\bb{p}$, following the guidelines of \citep{DodelsonSchneider13}. We consider three cases:
\begin{itemize}
\item If the true data covariance $\bb{C}$ is known, the estimator (\ref{estimatorcovariance}) is biased, and the dominant contribution of the bias comes from the second order fluctuations in $\bbh{\Psi}$. Once the expectation values are taken as in (\ref{psifluctuations}), the bias sums up to

\begin{equation}
\label{dodelsonscaling}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}(1+B(N_b-N_p))
\end{equation}
%
This is the result obtained by \citep{DodelsonSchneider13}.
\item If we are not interested in measuring the location parameter $\bbh{p}$, but just the width of the parameter posterior, we can tempted to replace the true covariance $\bb{C}$ with its estimator $\bbh{C}$. If we do that, and we do not correct for the bias (\ref{psibias}), the parameter variance will have a contribution from both the second and first order fluctuations in $\bbh{\Psi}$, which now have a non zero expectation value. In this case the bias sums up to 
\begin{equation}
\label{mockscalinguncorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}(1-(1+N_b)/N_r+B(1+N_p))
\end{equation}
\item If we repeat the same exercise as above, but we correct for the bias in the precision matrix estimator, we are left with

\begin{equation}
\label{mockscalingcorrected}
\langle\h{\Sigma}_\bb{p}\rangle=\Sigma_\bb{p}(1+B(1+N_p))
\end{equation}

\end{itemize} 
%
In the three previous cases we only kept the terms of order $O(1/N_r)$. The last two examples have to be considered as academic exercises, as in a real experiment the data has to be used in the end to compute the posterior location parameter and the parameter covariance. With the $\bbh{d}$ Gaussianity assumption, the coefficient $B$ can be calculated exactly as \citep{DodelsonSchneider13,Taylor12}:
\begin{equation}
B = \frac{N_r-N_b-2}{(N_r-N_b-1)(N_r-N_b-4)}
\end{equation}
%
In the limit $N_r\gg N_b$ this leads to a degradation of order $D/N_r$ for the variance of each estimated parameter $\sigma^2_p=\Sigma_{pp}$. We test the scaling relation
\begin{equation}
\label{ourscaling}
\langle\h{\sigma}_p^2\rangle = \sigma^2_p(N_s)\left(1+\frac{D}{N_r}\right)
\end{equation}
%
on our simulations. We call $D$ the \textit{effective dimensionality} of the feature space which, under Gaussianity assumptions, is equal to $D=N_b-N_p$ in the case where the true covariance $\bb{C}$ is known, $D=N_p-N_b$ when the true precision matrix is replaced by its biased estimator (computed from the same simulation set) and $D=1+N_p$ when the $\bbh{\Psi}$ is corrected.

We compute the expectation values of $\h{\sigma}^2_p$ (equations (\ref{estimatorcovariance}),(\ref{ourscaling})) averaging over 100 random resamplings of our shear ensembles. For the true feature covariance matrix $\langle(\bbh{d}-\bb{d}_0)(\bbh{d}-\bb{d}_0)^T\rangle=\bb{C}$ we use the estimated covariance from a grand ensemble built with the union of all the ensembles with different $N_s$.

The true parameter variance $\sigma^2_p(N_s)$ in principle depends on the number of independent $N$--body simulations $N_s$, which appears in the randomization procedure we describe in \S~\ref{shearsim}. The reason for this can be found in that, if $N_s$ is not big enough, the different shear realizations cannot be all independent, and hence the true variance $\sigma^2_p(N_s\rightarrow\infty)$ cannot be recovered in the limit $N_r\rightarrow\infty$. In particular there could be some residual error in the expectation value of (\ref{estimatorcovariance}) even if $N_r$ is big. In the next paragraph we outline our main findings.

@@

\begin{figure*}
\includegraphics[scale=0.4]{Figures/ps_pdf.eps}
\caption{PDF of the $\kappa$ power spectrum,$\mathcal{L}(P_l^{\kappa\kappa})$, at four selected multipoles $l=115,344,1349,5283$, for different shear ensembles constructed with $N_s=$1 (black), 2 (blue), 5 (green), 50 (red), 100 (purple). The thick black lines correspond to an ensemble generated with $N_s=1$ and $10^5$ mock realizations.}
\label{ps_pdf}
\end{figure*}

@@

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_nr.eps}
\caption{Bias in the $w$ variance, $\langle\h{\sigma}^2_w\rangle$, as a function of the number of mock realizations $N_r$ used in the feature covariance estimator (\ref{covest}). Results for different mock ensembles generated with $N_s=$1 (black), 2 (blue), 5 (green), 50 (red), 100 (purple) are shown. The figure shows both the measured trends (solid lines) and their best fits according to equation (\ref{ourscaling}). The thick black line correspond to an ensemble generated with $N_s=1$ and $10^5$ mock realizations. }
\label{wvar_nr}
\end{figure}

\begin{figure}
\includegraphics[scale=0.3]{Figures/scaling_ns.eps}
\caption{Variance on $w$ in the limit $N_r\rightarrow\infty$, varying the mock ensemble used to compute the covariance estimator (\ref{covest}). We show the dependence of $\sigma_0^2(N_s)$ in units of the mean over $N_s$, for the power spectrum logarithmically binned (black), the power spectrum linearly binned (red) and the peak counts (green).}
\label{wvar_ns}
\end{figure}

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\toprule
\textbf{Feature} &  \textbf{Specifications} & $N_b$ &  \textbf{Symbol} & \textbf{Color} \\ \hline \hline
\midrule
Power Spectrum, log binning  & $l \in [100,800] $ & 8 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [1000,6000] $ & 7 & $\times$ & black  \\ 
Power Spectrum, log binning  & $l \in [100,6000] $ & 15 & $\times$ & black  \\
Power Spectrum, linear binning  & $l \in [100,2000] $ & 15 & \textcolor{red}{$+$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [2500,4500] $ & 15 & \textcolor{red}{$\times$} & \textcolor{red}{red}  \\
Power Spectrum, linear binning  & $l \in [100,4500] $ & 30 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\ 
Power Spectrum, linear binning  & $l \in [100,6000] $ & 39 & \textcolor{red}{$\bullet$} & \textcolor{red}{red}  \\ \hline
Low peaks  & $\kappa_0 \in [-0.06,0.09] $ & 15 & \textcolor{OliveGreen}{$+$} & \textcolor{OliveGreen}{green}  \\ 
Intermediate peaks  & $\kappa_0 \in [0.1,0.27] $ & 15 & \textcolor{OliveGreen}{$\bigstar$} & \textcolor{OliveGreen}{green}  \\ 
High peaks  & $\kappa_0 \in [0.28,0.45] $ & 15 & \textcolor{OliveGreen}{$\diamond$} & \textcolor{OliveGreen}{green}  \\
Low+Intermediate peaks  & $\kappa_0 \in [-0.06,0.27] $ & 30 & \textcolor{OliveGreen}{$\times$} & \textcolor{OliveGreen}{green}  \\
Intermediate+High peaks  & $\kappa_0 \in [0.1,0.45] $ & 30 & \textcolor{OliveGreen}{$\bullet$} & \textcolor{OliveGreen}{green}  \\
All peaks  & $\kappa_0 \in [-0.06,0.45] $ & 45 & \textcolor{OliveGreen}{$\blacksquare$} & \textcolor{OliveGreen}{green}  \\ \hline
\bottomrule
\end{tabular}
\end{center}
\caption{Catalog of feature types used in this work, along with the chosen number of bands and the plot legends for Figure \ref{effectivenb}.}
\label{featuretable}
\end{table*}

\begin{figure*}
\includegraphics[scale=0.3]{Figures/effective_nb.eps}
\includegraphics[scale=0.3]{Figures/effective_nb_fake_uncorrected.eps}
\includegraphics[scale=0.3]{Figures/effective_nb_fake_corrected.eps}
\caption{Effective dimesionality of the feature space, $D$, as a function of the number of bands $N_b$, for the features outlined in Table \ref{featuretable}. The dimensionality $D$ is plotted in units of its predicted value from equations (\ref{dodelsonscaling}) in the left panel, equation (\ref{mockscalinguncorrected}) in the center panel and equation (\ref{mockscalingcorrected}) in the right panel.}
\label{effectivenb}
\end{figure*}

In this section we outline the main results of this work. We show the qualitative behavior of the $P^{\kappa\kappa}$ probability distribution function (PDF) in ensembles built with different $N_s$. We also show how the forecasts on the Dark Energy equation of state, $w$, depend on $N_s$ and $N_r$. Figure \ref{ps_pdf} shows the power spectrum PDF at four selected multipoles, while Figure \ref{ps_var} shows the variance in each multipole. This result is compared to the one that we would obtain assuming that the convergence is a Gaussian random field
\begin{equation}
\label{gaussianvar}
\mathrm{Var}(P^{\kappa\kappa}(l)) = \frac{P^{\kappa\kappa,2}(l)}{N(l)}
\end{equation}
%
where $N(l)$ is the number of modes used to estimate the power spectrum at $l$. $N_l$ is measured from the Fourier trasformed image grid. Figure \ref{wvar_nr} show the scaling of the expectation value of $\h{\sigma}_w^2$ with the number of mocks $N_r$ using the power spectrum and compare it with the known result obtained in \citep{DodelsonSchneider13} and expressed in equation (\ref{ourscaling}). Figure \ref{wvar_ns} shows the effect of varying $N_s$ on the $w$ constraint. Finally Figure \ref{effectivenb} shows how the effective feature dimensionality $D$ depends on the number of bands $N_b$, for a variety of features and binning choices outlined in Table \ref{featuretable}.

@@

In this section we discuss our main findings. Figure \ref{ps_pdf} shows that, although different choices of $N_s$ do not seem to affect the power spectrum PDF at large scales (top two panels), there are some qualitative differences on small scales (bottom two panels), on which a shear ensemble built with a small $N_s$ does not have the same statistical behavior of ensembles built with large $N_s$. Looking at the black curves, we see that the $N_s=1$ ensembles exibit some stochastic fluctuations in where the PDF peaks. We need as small as $N_s=2$ to recover the right location parameter for the power spectrum on small scales.

Figure \ref{ps_var} shows the variance of convergence power spectrum computed from different ensembles, in units of the Gaussian expectation. We find that, even with $N_s=1$, our results are in good agreement with the ones obtained by \citep{Sato12}, which used $N_s=400$. This fact by itself is not sufficient to conclude that $N_s$ does not have an effect on parameter inferences, since these depend on the cross band covariances.

In Figure \ref{wvar_nr} we examine how the $w$ constraint depends on the number of mocks used to estimate the covariance, and we find good agreement with (\ref{ourscaling}). In particular we can see that one $N$--body simulation is enough to construct an ensemble of $O(10^5)$ mutually independent mock power spectra, for which the $w$ constraint scales like $1/N_r$ all the way to $N_r=10^5$.

Figure \ref{wvar_ns} shows the dependence of the dark energy constraint with $N_s$. We find that, in the range $N_s\in[1,200]$ the $w$ inferred variance $\sigma_0^2$ fluctuates stochastically and does not show an appreciable trend with $N_s$.

We determine the effective dimensionality of the feature space, $D$, with a linear regression using equation (\ref{ourscaling}), for the set of features outlined in Table \ref{featuretable}. The results shown in Figure \ref{effectivenb}, however, show that the feature dimensionality $D$ grows faster than the linear $O(N_b)$ optimistic case predicted by \citep{DodelsonSchneider13}. When we estimate the real data covariance $\bb{C}$ from the same simulation set used to measure $\bbh{C}$, the effective dimensionality $D$ shows the expected behavior. More in detail, the toy models (\ref{mockscalinguncorrected}) and (\ref{mockscalingcorrected}) tell us that when we do not correct for the bias in the $\bbh{\Psi}$ estimator the effective dimensionality is negative (assuming $N_b>N_p$) and the parameter variance is underestimated when $N_r$ is not large enough. This explains what \citep{MinkPetri} observe in their Figures 7,8.
